\LARGE \bf ``Those who can't do, teach'': a Teachable Robotic Agent for Children Learning Handwriting

### 2014-09-19 main.tex
Full pass on the text

% * <deanna.m.hood@gmail.com> 2014-09-05T11:44:54.921Z:
%
% Ana Paiva is a reviewer....
%

@@

By leveraging simulated handwriting on a synchronised tablet display, a Nao
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories are generated, and allow
the robot to draw purposefully deformed letters. Learning how to write
characters well is then achieved by incorporating feedback from user
demonstrations, which allows the system to learn the optimal parameters for the
appropriate shape models.

Preliminary in situ studies have been conducted with primary school classes to obtain
insight into children's use of the novel system. 
%The technical components of the system have been validated, in that children 
Children aged 6-8 were successfully 1) led to believe that the robot was
actually writing on the tablet, and 2) able to improve the robot's writing to a
level which they were satisfied with. Furthermore, the system represents a
significant step towards an innovative use for robotics which addresses a
widespread and socially meaningful challenge in education, and provides context
for robotics in education beyond the well-explored use of teaching coding- and
robotics-related subjects.

@@

The \emph{learning by teaching} paradigm, which engages the target student in
the act of teaching another, has been shown to produce motivational,
meta-cognitive, and educational benefits in a range of disciplines
\cite{Rohrbeck2003}. An application which is yet to be explored, however, is
that of handwriting intervention. One reason why this has not been investigated
to date may be that in order to engage a child in the learning by teaching
paradigm, there must be an appropriately unskilled peer for them to tutor. In
the context where the target child is the lowest performer in their class, this
may present a logistical constraint. In some cases, it may be appropriate for a
peer or teacher to simulate a na\"ive learner for the target child to teach,
however for an activity where one's skill level is visually evident - as is the
case in handwriting - this acting is likely to be eventually detected. As such,
there is motivation for the use of a teachable agent which can be configured for
a variety of skill levels, and for which children do not have preconceptions
about handwriting ability.

@@

There are three main points which have motivated the use of a robot as the
teachable agent in the presented work:

@@

    \item observing a teachable agent respond to what it has been taught
        is a significant step in the learning by teaching
        paradigm \cite{Okita2006}, and therefore it may be significant for the agent to respond
        with a writing \emph{process}, rather than simply a product, and

@@

Consequently, the learning agent developed has been embodied in a commercially
affordable Nao humanoid robot. Its writing capabilities have been implemented by
leveraging synchronisation with a tablet display to simulate fine motor skills
otherwise infeasible for the robot. 
% * <deanna.m.hood@gmail.com> 2014-09-08T11:08:36.426Z:
%
% linking sentence...
%

Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a teachable
agent in the context of handwriting.  Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
novel approach for achieving simulated fine motor skills on commercially
affordable humanoid robots.  Section \ref{sec:experiment} explores the novel% * <deanna.m.hood@gmail.com> 2014-09-08T11:09:28.480Z:
%
% use a different word
%
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education.

@@

For the development of a reinforcement learning algorithm in the context of
handwriting, a parameterisation of letters and their deformities has been sought
to generate the necessary letter models. Depending on the parameters input to
the letter models, different quality shapes can be generated. The system can
therefore be configured to improve its writing by modifying the parameters used
to generate the letters, based on feedback from the reinforcement learning
partner. To this end, a method which can generate a model from real-world data
has been employed in order to capture realistic variations in letters.

@@

PCA has been performed on a set of letter paths captured from a digital pen,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 adult users). While it may be appropriate to
identify the location of the most salient features of the shapes in future work,
the features are currently taken as $n$ uniformly spaced points along the shape
path as salient features are not guaranteed from children's handwriting. The
points are arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path.

@@

In this work, the observation shapes are normalized such that they have a shape
centre of ${\bf 0}$ and a unit maximum dimension.

The projection from the original $2n$-dimensional feature space $(n=70)$ to an
$N$-dimensional space is found as in (\ref{eq:projection}), where ${\bf p}$
contains the coordinates in the $N$-dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. As such,
each shape may be approximated by the mean shape plus a sum of the top $N$
eigenvectors, weighted by the parameter vector $p$.

@@

For the dataset presented in Figure \ref{fig:deviations_sPrint}, the
eigenvectors associated with the 5 largest eigenvalues explain 90.0\% of the
variance in the dataset, illustrating the capability of the statistical shape
modelling approach to produce compact parameterisation of shapes.

Clusters within paths of a particular letter have been identified using K-means
clustering to group different styles of writing the letter (\emph{allographs}),
such that the parameters in the resulting model are only representative of the
specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%
PCA has been performed on all of the paths of a particular allograph in the
dataset at a time, to reduce the $2n$-dimensional space to one of a desired
number of dimensions.

@@

\begin{figure}[thpb]
\centering
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.05\textwidth]{figures/sPrint_mean.png}
}}
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.35\textwidth]{figures/sPrint_top3.png}
}}
\caption[The mean shape and the effect of varying the first three parameters of
the shape model derived from PCA from the dataset of print `s'
shapes.]{\label{fig:deviations_sPrint}The mean shape (left) and the effect
    (right) of varying the first three parameters (each row) of a shape model.
    The presented model has been derived from PCA on the first cluster of `s'
    shapes, interpreted as the print allograph. The extent to which each
parameter is varied is dependent on the eigenvector $\lambda$ corresponding to
the parameter's eigenvector. The percentage of the total variance in the dataset
explained by each parameter is shown to the left of the corresponding row. }

@@

Interestingly, although the parameters are the result of an \emph{unsupervised}
statistical shape analysis, they still represent variations which could have
been intuitively identified by a manual parameterisation. For example, for the
model shown in Figure \ref{fig:deviations_sPrint}, the parameters may represent
the width of the top half of the letter compared to the bottom half, the width
of the overall shape, etc. Consequently, the shape modelling process has
produced the desired outcomes of parameterising the shapes in ways which are
able to be used to deform them meaningfully.

@@

After determining the appropriate number of dimensions ($N$) for the shape
models, new letters can be generated by varying the parameter values in this
$N$-dimensional space in accordance with (\ref{eq:projection}). By choosing
parameter values which lie within the observed range in the dataset, it is
possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of reasonable,
adult-written letters, and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' - made from 120 observations each, with 70 sample points -
by selecting random values for the first 5 parameters from a distribution with
standard deviation of $3\sqrt\lambda_i$ rather than the $\sqrt\lambda_i$
standard deviation observed in the dataset.

\begin{figure}[thpb]
\centering
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.08\textwidth]{figures/e1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.05\textwidth]{figures/e4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.08\textwidth]{figures/e3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.08\textwidth]{figures/e5.png}}}\\
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.12\textwidth]{figures/g1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.12\textwidth]{figures/g3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.12\textwidth]{figures/g4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.12\textwidth]{figures/g5.png}}}
\caption[Sample letters generated from the PCA shape model on 120 `e' and `g'
paths.]{\label{fig:sampleLetters}Sample letters generated from the PCA shape
model on 120 `e' (top) and `g' paths (bottom), generated randomly from
parameters with $3\times$ the standard deviation observed in the dataset.}

@@

In \cite{Chandra2013}, it was found that children aged 4-6 years participating% * <deanna.m.hood@gmail.com> 2014-09-05T11:43:42.851Z:
%
% this is not peer-reviewed!
%
in a handwriting peer tutoring pilot study most often made mistakes classified
as inappropriate \emph{internal proportions} or \emph{global deformations}. It
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Using a database of children's letters when
available may yield potential for parameterising other mistakes made by
children, such as merging strokes or decomposing a shape into multiple subparts
inappropriately. Nevertheless, sufficient capabilities for the shape model to
generate `poor' letters which are appropriate for simulating a na\"ive learning
in the context of handwriting have been demonstrated, using the shape model
generated with PCA from a dataset of only adults' writing.

@@

It was shown in Section \ref{sec:writingGeneration} how the statistical shape
model of a particular letter may convert a set of parameters into a generated
letter of the same allograph. Similarly, the model may also be used to convert a
letter into its parameters, given the model. The parameters of user-drawn
letters may therefore be used in order to incorporate the capabilities for the
learning algorithm to adapt to the user's feedback via demonstration.

@@

The parameters which are determined for a shape in (\ref{eq:param}) will not
reconstruct the shape exactly if $N<2n$, but rather will reconstruct the closest
point in the $N$-dimensional space which the eigenvectors span. $N=10$ has been
used for the work presented.

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration. In the results presented in this work, the
update equation shown in \ref{eq:update} has been used, where ${\bf p}$ is the
learned parameter vector at time step $k$, and $\alpha$ is the learning rate,
between 0 and 1.

@@

It is possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, may produce parameters ${\bf p}^{(k+1)}$
by which yield an unacceptable shape. The proposed method for adapting to user
demonstration would not get stuck at such poor shapes, however: with further
demonstrations of the same letter, the system would eventually approach the
demonstration shape. As a result, the event of poor-looking intermediate letters
would not limit the interaction later proposed in Section \ref{sec:experiment}
in a technical sense, but it may influence the user's perception of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

Figure \ref{fig:demonstrationShapes2} illustrates the response of the system to
demonstrations from a child for the letter `u' using a learning rate of
$\alpha=1/2$.

\begin{figure}[thpb]
    \centering
    \includegraphics[width=0.3\textwidth]{figures/u_kids.png}
    \caption{\label{fig:demonstrationShapes2}Example of the learning algorithm
    responding to user demonstration of shapes for the letter `u.'}
\end{figure}

% * <deanna.m.hood@gmail.com> 2014-09-04T09:19:16.057Z:
%
% improve img
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An Embodied Handwriting Learning Agent with the Nao Humanoid Robot}
\label{sec:robotWriting}

In order to develop a teachable agent that is appropriate for engaging a child
through the learning by teaching paradigm, 
%- that is, that it is an embodied agent, capable of eliciting motor mimicry,
%and capable of causing its teacher to reflect on the process that has been used
%to respond to handwriting feedback -
capabilities for the robot to engage in handwriting have been established.

The Nao humanoid robot, which has been developed by Aldebaran-Robotics and
designed purposely to look approachable \cite{Gouaillier2008}, has been used for
this work. It is a commercially affordable biped robot, 58cm tall, with 25
degrees of freedom, two cameras, speech capabilities and the ability to
autonomously execute a range of tasks.

%Because of the challenges in getting the Nao to write trajectories smoothly
%because of its inherent ability to only reach discrete points in space, 
A fine level of control over what the robot is writing is necessary in the
proposed application of a learning agent in the context of handwriting. Because
of the limited fine motor skills possible with such an affordable robot, in
addition to the absence of force feedback and other technical necessities, the
Nao has been configured to use simulated handwriting with a synchronised tablet
to achieve this level of control. The necessary components of implementing the
proposed teachable robotic agent are, therefore:

\begin{enumerate}
    \item developing capabilities for the robot to trace handwriting,

@@

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child user in the action. Aldebaran's NaoQi API
has been used for the inverse kinematics of the trajectory following. The Robot
Operating System (ROS)\footnote{The ROS stack for Nao is available at:
\url{http://wiki.ros.org/nao_robot}} has been used for integration of the Nao
with external reference frames using the $tf$ transformation library
\cite{Foote2013}, such as the tablet's location.

@@

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect (inhuman) writing style are mitigated. % * <deanna.m.hood@gmail.com> 2014-09-04T09:59:44.707Z:
%
% is this point significant enough?
%

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.9\linewidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the writing of a word with
    its finger. The motion of the robot is synchronised with the display of the
    tablet, communicating over ROS.}%

   \label{fig:naoWriting}
\end{figure}

Therefore, the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education.}. As
interacting with a tablet with one's finger is not uncommon, this may aid the
acceptance of the writing style by users.

@@

An Android application has been developed to receive the trajectory message over
a ROS topic and add each point of the trajectory one by one as a frame of an
animation. Synchronisation between the tablet and the robot has been achieved by:

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/chilitagDetection_cameraOverlay.png}
    \caption{\label{fig:tabletDetection}Detection of the tablet using a fiducial
    marker to represent the origin of the writing surface frame. The robot's
    camera image is on the left (the text trajectory overlay is also visible).}

\end{figure}% * <deanna.m.hood@gmail.com> 2014-09-04T11:21:59.328Z:
%
% this image is pretty bad, and doesn't add much
%

A demonstration of the robot executing a trajectory in synch with the tablet is
shown in Figure \ref{fig:naoWriting}\footnote{See
\url{https://www.youtube.com/watch?v=2qWFSJRxCU0} for a video of the
demonstration. }.
% * <deanna.m.hood@gmail.com> 2014-09-04T11:40:30.176Z:
%
% this sounds dumb on paper.....
%

@@

The fusion of the embodied handwriting agent developed with the handwriting
learning algorithm presented in Section \ref{sec:learningAlgorithm} involves the
integration of three components: the robot, the tablet, and a central controller
(Figure~\ref{fig:archi}).  The robot and Android tablet application present the
writing process/result to the user, as in Section \ref{sec:tabletSynch}.
Furthermore, the tablet application is extended to act as the primary medium for
capturing user interaction, and publishes the user's demonstrations when user is
satisfied with their writing. The user demonstrations are received by the
interaction controller, running on a desktop computer. It is responsible for
getting the Nao to prompt and respond appropriately to feedback received, using
a finite state machine to manage the interaction stage and various system
inputs. It also manages the learning algorithm, by inferring the letter which
user demonstrations are for, based on their position on the tablet, and
publishes the shapes to be written which are generated by the learning algorithm
in response to the feedback.

@@

    \node [cmpt, fill=Turquoise!50, below=of tactile] (shape) {shape position \\ manager};
    \node [cmpt, fill=Turquoise!50, left=of shape] (learner) {letter \\ learner};
    \node [cmpt, fill=Turquoise!50, left=of learner] (card) {card \\ detector};

@@

\caption{Overview of the system. The top row runs on the tablet, the middle row on the central controller and the bottom row on the robot.}

@@

In addition to constituting a technically novel system, the presented teachable
robotic agent represents a tool which may be used for investigating social and
pedagogical research questions. One such question is what impact the addition of
such a teachable robotic agent would have on the outcomes of a typical
handwriting intervention.

@@

    \item The robot responds to the word request verbally, and writes the
        letters by pointing at the tablet and imitating tracing the letters. The
        tablet, placed in front of the robot with a vertical orientation,
        displays the letters in synchronisation with the robot's movements.

@@

    \item The controller produces an adapted letter for the robot to write in
        response to the participant's feedback, and the interaction iterates,
        with participants taking turns to interact with the robot if necessary.
        When the child is satisfied with the robot's performance on all letters,
        they may use the ``test'' card and an additional word on which they will
        verbally evaluate the robot's performance.

>>>>

@@

By leveraging simulated handwriting on a synchronised tablet display, a {\sc nao}
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories have been generated, and allow
the robot to draw purposefully deformed letters. Learning how to write
characters well is then achieved by incorporating feedback from user
demonstrations, which allows the system to learn the optimal parameters for the
appropriate shape models.

Preliminary in situ studies have been conducted with primary school classes to obtain
insight into children's use of the novel system. 
%The technical components of the system have been validated, in that children 
Children aged 6-8 were successfully 1) led to believe that the robot was
actually writing on the tablet, and 2) able to improve the robot's writing to a
level which they were satisfied with.

Furthermore, the system represents a
significant step towards an innovative use for robotics which addresses a
widespread and socially meaningful challenge in education, and provides context
for robotics in education beyond the well-explored use of teaching coding- and
robotics-related subjects.

@@

The \emph{learning by teaching} paradigm, which engages the target student in
the act of teaching another, has been shown to produce motivational,
meta-cognitive, and educational benefits in a range of disciplines
\cite{Rohrbeck2003}. An application which is yet to be explored, however, is
that of handwriting intervention. One reason why this has not been investigated
to date may be that in order to engage a child in the learning by teaching
paradigm, there must be an appropriately unskilled peer for them to tutor. In
the context where the target child is the lowest performer in their class, this
may present a logistical constraint. In some cases, it may be appropriate for a
peer or teacher to simulate a na\"ive learner for the target child to teach,
however for an activity where one's skill level is visually evident -- as is the
case in handwriting -- this acting is likely to be eventually detected. As such,
there is motivation for the use of a teachable agent which can be configured for
a variety of skill levels, and for which children do not have preconceptions
about its handwriting ability.

@@

The teachable agent developed has the potential to be embedded as a simulated 
computer-based agent, however there are three main points which have motivated 
the use of a robot instead as the teachable agent in the presented work:

@@

    \item observing a teachable agent respond to what it has been taught
        is a significant step in the learning by teaching
        paradigm \cite{Okita2006}, and therefore it may be significant for the agent to respond
        with a physical writing \emph{process}, rather than simply a product, and

@@

Consequently, the learning agent developed has been embodied in a commercially
affordable {\sc nao} humanoid robot. Its writing capabilities have been implemented by
leveraging synchronisation with a tablet display to simulate fine motor skills
otherwise infeasible for the robot.

Within this article, Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a teachable
agent in the context of handwriting. Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
new approach for achieving simulated fine motor skills on commercially
affordable humanoid robots.  Section \ref{sec:experiment} explores the 
contributions made to the study of human-robot interaction, in discussing the
use of the first-of-its-kind system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education. Finally, Section \ref{sec:futureWork} addresses the challenges which are faced in extending this system to a level suitable for long-term studies, and Section \ref{sec:conclusions} concludes by reiterating the content and impact of the article's contributions.

@@

For the development of a reinforcement learning algorithm in the context of
handwriting, a parameterisation of letters and their deformities has been sought
which is capable of generating the necessary letter models. Depending on the parameters input to
the letter models, different quality shapes should be generated. The system can
therefore be configured to improve its writing by modifying the parameters used
to generate the letters, based on feedback from the reinforcement learning
partner. To this end, a method which can generate a model from real-world data
has been employed in order to capture realistic variations in letters.

@@

PCA has been performed on a set of letter paths captured from a digital pen,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 adult users). While it may be appropriate in future work to
identify the location of salient features of the shapes which are robust to unanticipated user input,
the features are currently taken as $n$ uniformly spaced points along the shape
path. The
points are arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path. The observation shapes are normalised such that they have a shape
centre of ${\bf 0}$ and a unit maximum dimension.

@@

The projection from the original $2n$-dimensional feature space $(n=70)$ to a reduced
$N$-dimensional space is found as in (\ref{eq:projection}), where ${\bf p}$
contains the coordinates in the $N$-dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. As such,
each shape may be approximated by the mean shape plus a sum of the top $N$
eigenvectors, weighted by the parameter vector ${\bf p}$.

@@

For the dataset presented in Figure \ref{fig:deviations_sPrint}, the
eigenvectors associated with the 3 largest eigenvalues explain 78.5\% of the
variance in the dataset, illustrating the capability of the statistical shape
modelling approach to produce compact parameterisation of shapes.

%Clusters within paths of a particular letter have been identified using K-means
%clustering to group different styles of writing the letter (\emph{allographs}),
%such that the parameters in the resulting model are only representative of the
%specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%
PCA has been performed on all of the paths of a particular allograph in the
dataset at a time, to reduce the $2n$-dimensional space to one of a desired
number of dimensions. $N=10$ has been used for the work presented.

@@

\begin{figure}[thpb]
\centering
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.05\textwidth]{figures/sPrint_mean.png}
}}
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.35\textwidth]{figures/sPrint_top3.png}
}}
\caption[The mean shape and the effect of varying the first three parameters of
the shape model derived from PCA from the dataset of print `s'
shapes.]{\label{fig:deviations_sPrint}The mean shape (left) and the effect
    (right) of varying the first three parameters (each row) of a shape model.
     The extent to which each
parameter is varied is dependent on the eigenvalue $\lambda$ corresponding to
the parameter's eigenvector. The percentage of the total variance in the dataset
explained by each parameter is shown to the left of the corresponding row. }

@@

Interestingly, although the parameters are the result of an \emph{unsupervised}
statistical shape analysis, they still represent variations which could have
been intuitively identified by a manual parameterisation. For example, for the
model shown in Figure \ref{fig:deviations_sPrint}, the parameters may represent
the height of the top half of the letter compared to the bottom half, the width
of the overall shape, etc. Consequently, the shape modelling process has
produced the desired outcomes of parameterising the shapes in ways which are
able to be used to deform them meaningfully.

@@

After determining the appropriate number of dimensions ($N$) for the shape
models, new letters can be generated by varying the parameter values in this
$N$-dimensional space in accordance with (\ref{eq:projection}). By choosing
parameter values which lie within the observed range in the dataset, it is
possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of reasonable,
adult-written letters, and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' by selecting random values for the first 5 parameters 
from a distribution with
standard deviation of $3\sqrt\lambda_i$ rather than the $\sqrt\lambda_i$
standard deviation observed in the dataset.

\begin{figure}[thpb]
\centering
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.04\textwidth]{figures/e1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.025\textwidth]{figures/e4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.04\textwidth]{figures/e3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.04\textwidth]{figures/e5.png}}}~~~~~
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.06\textwidth]{figures/g1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.06\textwidth]{figures/g3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.06\textwidth]{figures/g4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[height=0.06\textwidth]{figures/g5.png}}}
\caption[Sample letters generated from the PCA shape model on `e' and `g'
paths.]{\label{fig:sampleLetters}Sample letters generated from the PCA shape
model on `e' (left) and `g' paths (right), generated randomly from
parameters with $3\times$ the standard deviation observed in the dataset.}

@@

In \cite{Chandra2013}, it was found that children aged 4-6 years participating
in a handwriting peer tutoring pilot study most often made mistakes classified
as inappropriate \emph{internal proportions} or \emph{global deformations}. It
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Using a database of children's letters when
available may yield potential for parameterising other mistakes made by
children, such as merging strokes or decomposing a shape into multiple subparts
inappropriately. Nevertheless, sufficient capabilities for the shape model to
generate `poor' letters which are appropriate for simulating a na\"ive learning
in the context of handwriting have been demonstrated, using the shape model
generated from PCA on a dataset of only adults' writing.

@@

It was shown in Section \ref{sec:writingGeneration} how the statistical shape
model of a particular letter may convert a set of parameters into a generated
letter of the same allograph. Similarly, the model may also be used ``in reverse'' 
to convert a
letter into its parameters, given the model. The parameters of user-drawn
letters may therefore be used in order to incorporate the capabilities for the
learning algorithm to adapt to the user's feedback via demonstration.

@@

The parameters which are determined for a shape in (\ref{eq:param}) will not
reconstruct the shape exactly if $N<2n$, but rather will reconstruct the closest
point in the $N$-dimensional space which the eigenvectors span.

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration. In the results presented in this work, the
linear update equation shown in (\ref{eq:update}) has been used, where ${\bf p}$ is the
learned parameter vector at time step $k$, and $\alpha$ is the learning rate,
between 0 and 1.

@@

Figure \ref{fig:demonstrationShapes2} illustrates the response of the system to
demonstrations from a child for the letter `s' using a learning rate of
$\alpha=1/2$.

\begin{figure}[thpb]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/learningSdemo}
    \caption{\label{fig:demonstrationShapes2}Example of the learning algorithm
    responding to user demonstration of shapes for the letter `s' (demonstrations received from two 7-8 year-old children).}
\end{figure}

It is possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, produce parameters ${\bf p}^{(k+1)}$
which yield an unacceptable shape, especially if the demonstration shape is of a different style to that learnt at time $k$, as there are no restrictions imposed on parameter values. The proposed method for adapting to user
demonstration would not get stuck at such poor shapes, however: with further
demonstrations of the same letter, the system would eventually approach the
demonstration shape (Figure \ref{fig:eDemo}. As a result, the event of poor-looking 
intermediate letters
would not limit the interaction later proposed in Section \ref{sec:experiment}
in a technical sense, but it may influence the user's \emph{perception} of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

\begin{figure}[thpb]
    \centering
    \includegraphics[width=0.3\textwidth]{figures/learningEdemo}
    \caption{\label{fig:eDemo}Example of the learning algorithm
    responding to user demonstration of shapes for the letter `e', passing through a parameter state which yields a poor letter (demonstrations received from a 7-8 year-old child).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An Embodied Handwriting Learning Agent with the NAO Humanoid Robot}
\label{sec:robotWriting}

In order to develop a teachable agent that is appropriate for engaging a child
through the learning by teaching paradigm, 
%- that is, that it is an embodied agent, capable of eliciting motor mimicry,
%and capable of causing its teacher to reflect on the process that has been used
%to respond to handwriting feedback -
capabilities for the robot to engage in handwriting and interactive turn-taking have been established.

The {\sc nao} humanoid robot, which has been developed by Aldebaran Robotics and
designed purposely to look approachable \cite{Gouaillier2008}, has been used for
this work. It is a commercially affordable biped robot, 58cm tall, with 25
degrees of freedom, two cameras, speech capabilities and the ability to
autonomously execute a range of tasks.

%Because of the challenges in getting the {\sc nao} to write trajectories smoothly
%because of its inherent ability to only reach discrete points in space, 
A fine level of control over what the robot is writing is necessary in the
proposed application of a learning agent in the context of handwriting. Because
of the limited fine motor skills possible with such an affordable robot, in
addition to the absence of force feedback and other technical necessities, the
{\sc nao} has been configured to use simulated handwriting with a synchronised tablet
to achieve this level of control. The necessary components of implementing the
proposed teachable robotic agent are, therefore:

\begin{enumerate}
    \item developing capabilities for the robot to trace handwriting ``in the air'',

@@

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child user in the action. Aldebaran's NaoQi API
has been used for the inverse kinematics of the trajectory following. The Robot
Operating System (ROS)\footnote{The ROS stack for {\sc nao} is available at:
\url{http://wiki.ros.org/nao_robot}.} has been used for integration of the {\sc nao}
with external reference frames, such as the tablet's location, using the $tf$ transformation library \cite{Foote2013}.

@@

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect writing form (\eg pen grip) are mitigated.

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.8\linewidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the writing of a word with
    its finger. The motion of the robot is synchronised with the display of the
    tablet, communicating over ROS.\protect\footnotemark}%

   \label{fig:naoWriting}
\end{figure}
\footnotetext{See
\url{https://www.youtube.com/watch?v=2qWFSJRxCU0} for a video of the synchronised writing 
demonstration. }

Therefore, the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education. The behaviour 
is hence not unfamiliar to children.}. As
interacting with a tablet with one's finger is not uncommon, this may aid the
acceptance of the writing style by users.

@@

An Android application has been developed to receive the trajectory message over
a ROS topic and display the trajectory as an
animation. Synchronisation between the tablet and the robot has been achieved by:

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.35\textwidth]{figures/chilitagDetection_cameraOverlay.png}
    \caption{\label{fig:tabletDetection}Detection of the tablet using a fiducial
    marker to represent the origin of the writing surface frame, visualised in RViz. The robot's
    camera image is on the left, with the text trajectory overlay visible.}

@@

The fusion of the embodied handwriting agent developed with the handwriting
learning algorithm presented in Section \ref{sec:learningAlgorithm} involves the
integration of three components: the robot, the tablet, and a central controller
(Figure~\ref{fig:archi}).  The robot and Android tablet application present the
writing process/result to the user, as in Section \ref{sec:tabletSynch}.
Furthermore, the tablet application has been extended to act as the primary medium for
capturing user interaction, and publishes the user's demonstrations when user is
satisfied with their writing. The user demonstrations are received by the
interaction controller, running on a desktop computer. It is responsible for
getting the {\sc nao} to prompt and respond appropriately to feedback received, using
a finite state machine to manage the interaction stage and various system
inputs. The controller also manages the learning algorithm, by inferring the letter which
user demonstrations are for based on their position on the tablet, and
publishes the shapes to be written which are generated by the learning algorithm
in response to the feedback.

@@

    \node [cmpt, fill=Green!50, below=of tactile] (shape) {shape position \\ manager};
    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};

@@

\caption{Overview of the system. Components in the top row run on the tablet, 
those in the middle row on the central controller, and those in the bottom row 
on the robot.}

@@

In addition to constituting a technically novel system, the presented teachable
robotic agent represents a tool which may be used for investigating social and
pedagogical research questions. For example, one such question is what impact the addition of
such a teachable robotic agent would have on the outcomes of a typical
handwriting intervention.

@@

    \item The robot responds to the word request verbally, and writes the
        letters by pointing at the tablet and imitating tracing the letters. The
        tablet, placed in front of the robot with a vertical, landscape orientation,
        displays the letters in synchronisation with the robot's movements.

@@

    \item The controller produces an adapted letter for the robot to write in
        response to the participant's feedback, and the interaction iterates,
        with participants taking turns to interact with the robot if necessary.
        When the participant(s) is/are satisfied with the robot's performance on 
	all letters,
        they may use the ``test'' card and an additional word for which they will
        verbally evaluate the robot's performance.
